---
title: "Developing Interactive Visualization Tools for Model-Supported R&D"
output: 
  beamer_presentation:
    theme: boxes
    fig_width: 4
    fig_height: 3
    includes:
      in_header: "mrgbeam.tex"
---

```{r,echo=FALSE,message=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE)
library(dplyr)
library(magrittr)
library(dmutate)
```



Outline
===========

- Where we're heading: moxifloxacin PK dashboard app 

- Some `mrgsolve`

- Some `Shiny`

- Create a simple app to explore the moxi model

- Review the moxi dashboard app

- Deploy the moxi app on Metworx Envision

- Look at other apps for clinical decision-making

- Metworx Envison hands-on


Moxifloxacin dashboard app
==================================
- https://metrumrg.shinyapps.io/moxi
- [Moxifloxacin PK in diabetic foot infection](http://repository.ddmore.eu/model/DDMODEL00000034)
- Proposed dose: 400 mg IV Q24h


Decision questions
==================================
- Under the proposed dose, what is the highest MIC we can expect to cover?
   - \redb{M}inimum \redb{I}nhibitory \redb{C}oncentration
   - Gram positive bacteria
   - Gram negative bacteria
- What is the expected efficacy?
  - AUC/MIC ratio
      - After first dose
      - After third dose
- How does patient body size affect expected efficacy?
  - Weight / BMI


To answer these questions, we need ...
============================================
- Moxifloxacin population PK model
    - Day 1 and day 3 AUC
    - Loading and maintenance doses
   - Fraction unbound
- Simulate populations
    - Different body-size distribution
- Summaries by several factors
    - Bacteria type
    - MIC
  
A dashboard gives access to decision makers
==========================================
- Real-time updates based on decision-maker input
- Let decision makers
    - interact with the model
    - explore the scenarios of their choosing
    - answer questions as the discussion develops
- Does not require any technical skill or modeling background



What we need for the dashboard
========================================
- Moxifloxacin PK model in agile format
    - Needs to be quick
    - Clean implementation
    - Accessible / portable
- User interface
    - Allow user input
    - Display simulation results
    - Easy to understand / navigate
- Some code that turns user input into desired results


mrgsolve in brief
=======================

```{r,message=FALSE}
library(mrgsolve)
mod <- mread_cache("irm2", modlib())
```

```{r}
mod
```

Simulate a dosing regimen
===============
```{r, fig.width=4,fig.height=3}
out <- 
  mod %>% 
  param(KOUT=0.2,IC50=2) %>%
  ev(amt=400,ii=24, addl=4) %>% 
  mrgsim(end=240, delta=0.1)
```

***
```{r,fig.align="center",fig.width=4}
plot(out,EV1+CP+RESP~.)
```


Population simulation
===============
```{r}
mod <- mread("popex", modlib())
```

```{r}
cov1 <- covset(WT~rnorm(80,30),SEX~rbinomial(0.4))
idata <- data_frame(ID=1:25) %>% mutate_random(cov1) 

head(idata,n=3)

out <- mod %>% ev(amt=100) %>% mrgsim(end=96,idata=idata)
```

***
```{r,fig.width=3,fig.align="center"}
plot(out,DV~.)
```

Get started with mrgsolve
==============================
- [mrgsolve.github.io](https://mrgsolve.github.io)


Moxifloxacin model
=======================
- Let's do side-by-side conversion from NONMEM control stream
- Model downloaded from DDMoRe model repository
    - http://repository.ddmore.eu/model/DDMODEL00000034
- First, do a one-to-one conversion
    - Model amounts computed in closed form (like ADVAN2/ADVAN4)
- Then, code an ODE-based version
    - This let's us get the day 1 and 3 `AUC` values we need for the dashboard
 

$PROBLEM
=============

\dsbb{NONMEM}
```{r,eval=FALSE}
;; 1. Based on: run021
;; 2. Description: PK of moxifloxacin in patients with diabetic foot infections
;; x1. Author: Sebastian G. Wicha, Freie Universitaet Berlin, Germany
;; 3. Label:
$PROBLEM PK
```

\redb{mrgsolve}
```{r,eval=FALSE}
$PROB
# Wicha et al. 2015
- Moxifloxacin PK in diabetic foot infection
- Author: Sebastian G. Wicha, Freie Universitaet Berlin, Germany
- https://www.ncbi.nlm.nih.gov/pubmed/25600294
```

$INPUT
=======
\dsbb{NONMEM}
```{r,eval=FALSE}
$INPUT ID TIME DV AMT RATE MDV EVID CMT WGT IBW 
```

\redb{mrgsolve}
```{r,eval=FALSE}
$PARAM WGT = 70, IBW = 70, fu = 0.6
```


$DATA
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
$DATA MOX02_sim.csv ; IGNORE=#
```
\redb{mrgsolve}
```{r,eval=FALSE}
## Not required
```

- `mrgsolve` will work with your NONMEM data set
- Pass in as `R` data frame


$SUBROUTINES
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
$SUBROUTINES ADVAN3 TRANS4
```


\redb{mrgsolve}
```{r,eval=FALSE}
$PKMODEL cmt="CENT PERIPH"
```

$PK
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
$PK
TVCL  = THETA(1) 
ASCCL = TVCL * (IBW/70)**0.75
CL    = ASCCL * EXP(ETA(1))\
TVV1  = THETA(2) 
ASCV1 = TVV1 * (WGT/70)**1
V1    = ASCV1 * EXP(ETA(2))
Q  = THETA(3) 
TVV2 = THETA(4)
ASCV2 = TVV2 * (WGT/70)**1
V2    = ASCV2 
S1 = V1
```

***
\redb{mrgsolve}
```{r,eval=FALSE}
$MAIN
double TVCL  = THETA1;
double ASCCL = TVCL*pow(IBW/70,0.75);
double CL    = ASCCL*exp(ETA(1));

double TVV1  = THETA2;
double ASCV1 = TVV1*WGT/70.0;
double V1    = ASCV1*exp(ETA(2));

double Q = THETA3;

double TVV2  = THETA4;
double ASCV2 = TVV2*WGT/70.0;
double V2    = ASCV2;
```




$ERROR
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
IPRED = F
DEL   = 0
IF (IPRED.EQ.0) DEL=0.0001
W     = F
Y     = F+F*EPS(1)+EPS(2)
IRES  = DV-IPRED
IWRES = IRES/(W+DEL)
```

***
\redb{mrgsolve}
```{r,eval=FALSE}
$TABLE
capture IPRED = CENT/V1;
double  DV = IPRED*(1+EPS(1)) + EPS(2);

capture fIPRED = fu*IPRED;
capture fDV = fu*DV;

$CAPTURE DV
```


$THETA, $OMEGA, $SIGMA
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
$THETA
1.21E+01 ; CL
6.81E+01 ; V1
2.03E+01 ; Q
4.46E+01 ; V2

$OMEGA
6.33E-02        ; CL
7.25E-02        ; V1

$SIGMA
3.40E-02 ; Prop 
4.65E-03 ; Add
```

***
\redb{mrgsolve}
```{r,eval=FALSE}
$THETA
1.21E+01
6.81E+01
2.03E+01
4.46E+01

$OMEGA
6.33E-02
7.25E-02

$SIGMA
3.40E-02
4.65E-03
```

No equivalent
=======


\dsbb{NONMEM}
```{r,eval=FALSE}
$EST METHOD=1 INTER MAXEVAL=0 
NOABORT SIG=3 PRINT=1 POSTHOC
;$COV
```

\redb{mrgsolve}
```{r,eval=FALSE}
$SET delta=0.1, end=24*3
```

- Easy to simulate at the times you want observations
- `mrgsolve` itself doesn't not generate estimates; but
it may be used as a part of optimization process in `R`

Let's expand the mrgsolve model
========
- Differential equations
- Calculate AUC0-24 and AUC48-72


Develop R code for moxi sims
===============================
- Need cohort with 1000 patients
- Covariates
    - Weight, height,SEX (IBW,BMI)
- Dosing
    - 400 mg IV Q24h
    - Option for loading dose (hypothetical)
- Simulate out to day 3

Introduce dmutate package
=========
- [https://cran.r-project.org/package=dmutate](https://cran.r-project.org/package=dmutate)
- Quickly simulate variates
     - any distribution at any level 
     - bounds 
- Specified as formula
- Collect in sets

dmutate example
=====================
```{r}
library(dmutate)
muwt <- 90
sdwt <- 90
pfe <- 0.65
```

```{r}
cov1 <- covset(WT[40,140] ~ rnorm(muwt,sdwt)|ID,
               SEX ~ rbinomial(pfe)|ID)
```
```{r}
cov1
```

***
```{r}
set.seed(1133)
data_frame(ID=1:25) %>% mutate_random(cov1)
```

***
```{r}
data(exTheoph)

exTheoph %<>% mutate_random(cov1)

head(exTheoph)

```


Shiny
====================
- A huge topic; we will just scratch the surface
- Developed by Rstudio + lots of contributors
- The place to start: http://shiny.rstudio.com/
    - Tutorials, galleries, articles, example code


Elements of Shiny
====================
- Two main elements to shiny app
    - User interface (UI, `ui.R`)
    - Server (`server.R`)
- User interface
    - Data input (widgets: text, numeric, select, checkbox)
    - Return data (plots, tables, text)
    - Layout (fluidPage, tabSet)
- Server
    - All computations
    - Generate outputs to be displayed in UI
    - Largely plain old `R` code written in server environment
- Reactivity

Explore Shiny with a flexdashboard
====================================
- Interactive Rmarkdown document
    - Define some inputs 
    - Return some outputs
- flexdashboard puts it all together for you
    - Widgets (input)
    - Render (output)
    - Reactivity
    - Control reactivity with action button


Recommend shinydashboard as environment for your app
===============================
- https://rstudio.github.io/shinydashboard/
- Similar concept to flexdashboard
- Menu bar on the left
    - Add submenus
- Main app space on the right
    - Arrange with boxes
    - Add tabsets


Let's take a look at the Moxifloxacin code
============================================
- `./moxi2/ui.R` and `./moxi2/server.R`







